---
title: "FinalResults_05-Prediction_Evaluation-02-K-NearestNeighbours"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Predicting Online News Popularity based on Pre-publication Features

Final Results and Code


Step 5 - Prediction & Step 6 - Performance Evaluation

Prediction: Classification models - Logistic Regression, K-Nearest Neighbours, Decision Trees, Random Forest, and Gradient Boosting Machine - are tested to find the algorithm with the best accuracy. 

Performance Evaluation: At this stage, the machine learning algorithms that have been run are evaluated using various methods. Several metrics are computed including Accuracy, Precision, Recall/Sensitivity, Specificity, and F1, depending on the Confusion Matrix. 

Note the results found are similar to what has been found in the literature.


K-Nearest Neighbours

```{r }
#Reload the data
train.set <- read.csv(file = "trainset.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE, 
                      na.strings = c("", "NA"))

valid.set <- read.csv(file = "validset.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE, 
                      na.strings = c("", "NA"))

test.set <- read.csv(file = "testset.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE, 
                     na.strings = c("", "NA"))

train <- read.csv(file = "train.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE, 
                  na.strings = c("", "NA"))

test <- read.csv(file = "test.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE, 
                 na.strings = c("", "NA"))

#Remove the index column
train.set <- train.set[2:28]
valid.set <- valid.set[2:28]
test.set <- test.set[2:28]
train <- train[2:28]
test <- test[2:28]
```

```{r }
#Change the categorical variables back to factor to conduct model training and testing
train.set$data_channel_is_entertainment <- as.factor(train.set$data_channel_is_entertainment)
train.set$data_channel_is_socmed <- as.factor(train.set$data_channel_is_socmed)
train.set$weekday_is_saturday <- as.factor(train.set$weekday_is_saturday)
train.set$weekday_is_sunday <- as.factor(train.set$weekday_is_sunday)
train.set$shares_cat <- as.factor(train.set$shares_cat)

valid.set$data_channel_is_entertainment <- as.factor(valid.set$data_channel_is_entertainment)
valid.set$data_channel_is_socmed <- as.factor(valid.set$data_channel_is_socmed)
valid.set$weekday_is_saturday <- as.factor(valid.set$weekday_is_saturday)
valid.set$weekday_is_sunday <- as.factor(valid.set$weekday_is_sunday)
valid.set$shares_cat <- as.factor(valid.set$shares_cat)

test.set$data_channel_is_entertainment <- as.factor(test.set$data_channel_is_entertainment)
test.set$data_channel_is_socmed <- as.factor(test.set$data_channel_is_socmed)
test.set$weekday_is_saturday <- as.factor(test.set$weekday_is_saturday)
test.set$weekday_is_sunday <- as.factor(test.set$weekday_is_sunday)
test.set$shares_cat <- as.factor(test.set$shares_cat)

train$data_channel_is_entertainment <- as.factor(train$data_channel_is_entertainment)
train$data_channel_is_socmed <- as.factor(train$data_channel_is_socmed)
train$weekday_is_saturday <- as.factor(train$weekday_is_saturday)
train$weekday_is_sunday <- as.factor(train$weekday_is_sunday)
train$shares_cat <- as.factor(train$shares_cat)

test$data_channel_is_entertainment <- as.factor(test$data_channel_is_entertainment)
test$data_channel_is_socmed <- as.factor(test$data_channel_is_socmed)
test$weekday_is_saturday <- as.factor(test$weekday_is_saturday)
test$weekday_is_sunday <- as.factor(test$weekday_is_sunday)
test$shares_cat <- as.factor(test$shares_cat)
```

```{r }
library("class")
#install.packages("gmodels")
library("gmodels")

#Remove the share categories column from the training and test datasets
train.set_new <- train.set[-27]
test.set_new <- test.set[-27]
valid.set_new <- valid.set[-27]

train_new <- train[-27]
test_new <- test[-27]

#Store the labels
train.set_labels <- train.set$shares_cat
test.set_labels <- test.set$shares_cat
valid.set_labels <- valid.set$shares_cat

train_labels <- train$shares_cat
test_labels <- test$shares_cat
```

```{r }
#Tune the model for different levels of k (number of neighbours)
library(caret)
set.seed(123)
knnfit <- train(x = train.set_new, y = train.set_labels, method = "knn", tuneLength = 10, 
                trControl = trainControl(method = "cv", search = "grid"))

#Print the training model to see the accuracy
print(knnfit)
```

```{r }
#Plot the results
plot(knnfit)
#The most accurate predictions are made at k = 23
```

```{r }
#Validate and test the model with k = 23
knn_prediction <- knn(train = valid.set_new, test = test.set_new, cl= valid.set_labels, k = 23)

#Confusion Matrix and Statistics
knn_ConfusionMatrix_stats = confusionMatrix(knn_prediction, test.set_labels, mode = "everything", positive = "1")
print(knn_ConfusionMatrix_stats)
```


Using 10-fold Cross-Validation

```{r }
#For k = 23, make a prediction on the test set.
knn_prediction2 <- knn(train = train_new, test = test_new, cl= train_labels, k = 23)

#Confusion Matrix and Statistics
library(caret)
knn_ConfusionMatrix_stats2 = confusionMatrix(knn_prediction2, test_labels, mode = "everything", positive = "1")
print(knn_ConfusionMatrix_stats2)
```
