---
title: "FinalResults_05-Prediction_Evaluation-03-DecisionTrees"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Predicting Online News Popularity based on Pre-publication Features

Final Results and Code


Step 5 - Prediction & Step 6 - Performance Evaluation

Prediction: Classification models - Logistic Regression, K-Nearest Neighbours, Decision Trees, Random Forest, and Gradient Boosting Machine - are tested to find
the algorithm with the best accuracy. 

Performance Evaluation: At this stage, the machine learning algorithms that have been run are evaluated using various methods. Several metrics are computed
including Accuracy, Precision, Recall/Sensitivity, Specificity, and F1, depending on the Confusion Matrix. 

Note the results found are similar to what has been found in the literature.


Decision Trees

```{r }
#Reload the data
train.set <- read.csv(file = "trainset.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE, 
                      na.strings = c("", "NA"))

valid.set <- read.csv(file = "validset.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE, 
                      na.strings = c("", "NA"))

test.set <- read.csv(file = "testset.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE, 
                     na.strings = c("", "NA"))

train <- read.csv(file = "train.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE, 
                  na.strings = c("", "NA"))

test <- read.csv(file = "test.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE, 
                 na.strings = c("", "NA"))

#Remove the index column
train.set <- train.set[2:28]
valid.set <- valid.set[2:28]
test.set <- test.set[2:28]
train <- train[2:28]
test <- test[2:28]
```

```{r }
#Change the categorical variables back to factor to conduct model training and testing
train.set$data_channel_is_entertainment <- as.factor(train.set$data_channel_is_entertainment)
train.set$data_channel_is_socmed <- as.factor(train.set$data_channel_is_socmed)
train.set$weekday_is_saturday <- as.factor(train.set$weekday_is_saturday)
train.set$weekday_is_sunday <- as.factor(train.set$weekday_is_sunday)
train.set$shares_cat <- as.factor(train.set$shares_cat)

valid.set$data_channel_is_entertainment <- as.factor(valid.set$data_channel_is_entertainment)
valid.set$data_channel_is_socmed <- as.factor(valid.set$data_channel_is_socmed)
valid.set$weekday_is_saturday <- as.factor(valid.set$weekday_is_saturday)
valid.set$weekday_is_sunday <- as.factor(valid.set$weekday_is_sunday)
valid.set$shares_cat <- as.factor(valid.set$shares_cat)

test.set$data_channel_is_entertainment <- as.factor(test.set$data_channel_is_entertainment)
test.set$data_channel_is_socmed <- as.factor(test.set$data_channel_is_socmed)
test.set$weekday_is_saturday <- as.factor(test.set$weekday_is_saturday)
test.set$weekday_is_sunday <- as.factor(test.set$weekday_is_sunday)
test.set$shares_cat <- as.factor(test.set$shares_cat)

train$data_channel_is_entertainment <- as.factor(train$data_channel_is_entertainment)
train$data_channel_is_socmed <- as.factor(train$data_channel_is_socmed)
train$weekday_is_saturday <- as.factor(train$weekday_is_saturday)
train$weekday_is_sunday <- as.factor(train$weekday_is_sunday)
train$shares_cat <- as.factor(train$shares_cat)

test$data_channel_is_entertainment <- as.factor(test$data_channel_is_entertainment)
test$data_channel_is_socmed <- as.factor(test$data_channel_is_socmed)
test$weekday_is_saturday <- as.factor(test$weekday_is_saturday)
test$weekday_is_sunday <- as.factor(test$weekday_is_sunday)
test$shares_cat <- as.factor(test$shares_cat)
```

```{r }
#install.packages("party")
library("party")

#Tune the model based on mincriterion - the minimum value of the test statistic (1 - p-value)
library(caret)
set.seed(123)
ctreefit <- train(shares_cat ~., data = train.set, method = "ctree", tuneLength = 6, 
                  trControl = trainControl(method = "cv", search = "grid"))

#Print the training model to see the accuracy
print(ctreefit)
```

```{r }
#Plot the results
plot(ctreefit)
```

The most accurate predictions were made with mincriterion = 0.794
Since mincriterion = 0.99 is very close in accuracy, this value may be chosen for the model since mincriterion is determined by the p-value.

```{r }
#Prune the tree changing the control parameter to mincriterion = 0.99
#Run the model on the validation set
set.seed(123)
ctree_pruned <- ctree(shares_cat ~ ., data = valid.set, controls = ctree_control(mincriterion = 0.99))

#Make a prediction on the test set
ctree_pruned_prediction <- predict(ctree_pruned, test.set)

#Confusion Matrix and Statistics
ctree_pruned_ConfusionMatrix_stats = confusionMatrix(ctree_pruned_prediction, test.set$shares_cat,
                                                     mode = "everything", positive = "1")
print(ctree_pruned_ConfusionMatrix_stats)
```


Using 10-fold cross-validation

```{r }
#Create a decision tree model on the training set using the best parameters
set.seed(123)
ctree_model <- ctree(shares_cat ~. , data = train, controls = ctree_control(mincriterion = 0.99))

#Print the decision tree
print(ctree_model)
```

At the root node of the tree is closeness to LDA topic 2. This variable will help decide the class of popularity.

Some important variables found to help determine whether an article will be popular are closeness to LDA topic 2 (statistic = 869.307), followed by whether the
data channel is Entertainment (statistic = 736.054), whether the article was published on Saturday (statistic = 279.203), minimum shares of the average keyword
(statistic = 216.583), whether the data channel is Social Media (statistic = 161.159), number of links (statistic = 149.883), whether the day the article was
posted is Sunday (statistic = 113.296), average shares of the best keyword (statistic = 92.725), number of images (statistic = 58.131), and closeness to LDA topic
0 (statistic = 54.745).

```{r }
#Make a prediction on the test set.
ctree_prediction <- predict(ctree_model, test)

#Confusion Matrix and Statistics
ctree_model_ConfusionMatrix_stats = confusionMatrix(ctree_prediction, test$shares_cat, mode = "everything", 
                                                    positive = "1")
print(ctree_model_ConfusionMatrix_stats)
```
